Inferring the distribution and rate of Earth-like exoplanets \\ from noisy individual-planet inferences

### 2014-05-19 document/ms.tex
a few words about the real rate

An impressive number of exoplanets have been discovered and it is of great
scientific interest to study what we can say about the exoplanet occurence
rate function given these discoveries.
Every catalog is censored by non-trivial selection effects and detection
efficiencies, and the measurements of the system properties (period, radius,
\etc) are noisy.
We present a general hierarchical probabilistic framework for making justified
inferences about the occurrence rate of exoplanets taking into account survey
completeness \emph{and observational uncertainties}.
By applying our method to synthetic catalogs, we demonstrate that it produces
more accurate estimates of the underlying occurence rate than the standard
inverse-detection-efficiency weighting procedure.
When applied to an existing catalog of small planet candidates around G dwarfs
(Petigura \etal\ 2013), our method confirms previous measurements of a
flattening radius distribution near Earth's radius.
Although no true extra-solar Earth analog is known, hundreds of planets have
been found around Sun-like stars that are either Earth-sized but on shorter
periods, or else on year-long orbits but somewhat larger.
Under strong assumptions, it has been shown that these populations permit an
uncertain extrapolated estimate of the probability or rate at which Sun-like
stars host true Earth analogs.
We relax these assumptions by modeling the occurence rate as a
multidimensional Gaussian process and find that the rate density of Earth
analogs is about two percent (per natural logarithmic bin in period and
radius) with large uncertainty.

@@

\section{Data and completeness function}

@@

\begin{figure}[p]
\begin{center}
\includegraphics[width=\textwidth]{figures/smooth/results.pdf}
\end{center}
\caption{%
Inferences about the occurence rate function based on the simulated catalog
\modela.
\emph{Center:} the points with error bars show the exoplanet candidates in the
simulated incomplete catalog, the contours show the survey completeness
function (\citealt{petigura}), and the grayscale shows the median posterior
occurence surface.
\emph{Top and left:} the red dashed line shows the \emph{true} distribution
that was used to generate the catalog, the points with error bars show the
results of the inverse-detection-efficiency procedure, and the histograms are
posterior samples from the marginalized occurence rate as inferred by our
method.
\figlabel{smooth-results}}
\end{figure}

\begin{figure}[p]
\begin{center}
\includegraphics[width=\textwidth]{figures/simulation/results.pdf}
\end{center}
\caption{%
The same as \fig{smooth-results} for \modelb.
\figlabel{simulation-results}}
\end{figure}

\section{Extrapolation to Earth}

@@

\begin{figure}[p]
\begin{center}
\includegraphics[width=0.6\textwidth]{figures/smooth/rate.pdf}
\end{center}
\caption{%
The extrapolated rate of Earth analogs \gammaearth\ as inferred by the
different techniques.
Applying the method used by \citet{petigura} gives a constraint indicated by
the vertical black line with error bars shown as dashed lines.
The histogram is the MCMC estimate of our posterior constraint on this rate
and the true value is indicated as the thick gray vertical line.
\figlabel{smooth-rate}}
\end{figure}

\begin{figure}[p]
\begin{center}
\includegraphics[width=0.6\textwidth]{figures/simulation/rate.pdf}
\end{center}
\caption{%
The same as \fig{smooth-rate} for \modelb.
\figlabel{simulation-rate}}
\end{figure}

\section{Real data}

>>>>

An impressive number of exoplanets have been discovered and it is of great
scientific interest to study what we can say about the exoplanet occurence
rate function given these discoveries.
Every catalog is censored by non-trivial selection effects and detection
efficiencies, and the measurements of the system properties (period, radius,
\etc) are noisy.
We present a general hierarchical probabilistic framework for making justified
inferences about the occurrence rate of exoplanets taking into account survey
completeness \emph{and observational uncertainties}.
By applying our method to synthetic catalogs, we demonstrate that it produces
more accurate estimates of the underlying occurence rate than the standard
inverse-detection-efficiency weighting procedure.
When applied to an existing catalog of small planet candidates around G dwarfs
(Petigura \etal\ 2013), our method confirms previous measurements of a
flattening radius distribution near Earth's radius.
Although no true extra-solar Earth analog is known, hundreds of planets have
been found around Sun-like stars that are either Earth-sized but on shorter
periods, or else on year-long orbits but somewhat larger.
Under strong assumptions, it has been shown that these populations permit an
uncertain extrapolated estimate of the probability or rate at which Sun-like
stars host true Earth analogs.
We relax these assumptions by modeling the occurence rate as a
multidimensional Gaussian process and find that the rate density of Earth
analogs is about two percent (per natural logarithmic bin in period and
radius) at low signal-to-noise.

@@

\section{Data and completeness function}
\sectlabel{data}

@@

\begin{figure}[p]
\begin{center}
\includegraphics[width=\textwidth]{figures/smooth/results.pdf}
\end{center}
\caption{%
{\bf Simulated data}.
Inferences about the occurence rate function based on the simulated catalog
\modela.
\emph{Center:} the points with error bars show the exoplanet candidates in the
simulated incomplete catalog, the contours show the survey completeness
function (\citealt{petigura}), and the grayscale shows the median posterior
occurence surface.
\emph{Top and left:} the red dashed line shows the \emph{true} distribution
that was used to generate the catalog, the points with error bars show the
results of the inverse-detection-efficiency procedure, and the histograms are
posterior samples from the marginalized occurence rate as inferred by our
method.
\figlabel{smooth-results}}
\end{figure}

\begin{figure}[p]
\begin{center}
\includegraphics[width=\textwidth]{figures/simulation/results.pdf}
\end{center}
\caption{%
{\bf Simulated data}.
The same as \fig{smooth-results} for \modelb.
\figlabel{simulation-results}}
\end{figure}

\section{Extrapolation to Earth}
\sectlabel{extrap}

@@

\begin{figure}[p]
\begin{center}
\includegraphics[width=0.6\textwidth]{figures/smooth/rate.pdf}
\end{center}
\caption{%
{\bf Simulated data}.
The extrapolated rate of Earth analogs \gammaearth\ as inferred by the
different techniques applied to the \modela\ simulation.
Applying the method used by \citet{petigura} gives a constraint indicated by
the vertical black line with error bars shown as dashed lines.
The histogram is the MCMC estimate of our posterior constraint on this rate
and the true value is indicated as the thick gray vertical line.
\figlabel{smooth-rate}}
\end{figure}

\begin{figure}[p]
\begin{center}
\includegraphics[width=0.6\textwidth]{figures/simulation/rate.pdf}
\end{center}
\caption{%
{\bf Simulated data}.
The same as \fig{smooth-rate} for \modelb.
\figlabel{simulation-rate}}
\end{figure}

\section{Results from real data}

Having developed this probabilistic framework for occurence rate inferences
and demonstrating that it produces reasonable results when applied to
simulated datasets, we now turn to real data.
As described in \sect{data}, we will use the catalog of small exoplanet
candidates orbiting Sun-like stars published by \citet{petigura}.
This is a great test case because those authors empirically measured the
detection efficiency of their pipeline as a function of the parameters of
interest.

\Fig{real-results} shows posterior samples from the inferred occurence rate
distribution as a function of period and radius conditioned on the catalog.
The marginalized distributions are qualitatively consistent with the occurence
rate functions measured by \citet{petigura}.
The MCMC chain used to generate this figure is available online at
\dfmtodo{ADD SOME URL}.

Our constraint on the rate of Earth analogs (as defined in \sect{extrap}) is
in tension---even though our result has large fractional uncertainty---with
the result from \citet{petigura}.
This is shown in \fig{real-rate} where we compare the marginalized posterior
PDF for \gammaearth\ to the published value and uncertainty.
Quantitatively, we find that the rate of Earth analogs---the average number of
Earth-sized planets orbiting a Sun-like on a year-long period---is
\begin{eqnarray}
\gammaearth &=& 0.017^{+0.018}_{-0.009} \quad.
\end{eqnarray}
Converted to these units, \citet{petigura} measured $0.119_{-0.046}^{+0.035}$
for the same quantity (indicated as the vertical lines in \fig{real-rate}).
